Rule Synthesis for Rule-based Intrusion Detectors

### 2020-06-02 paper/paper.tex
pass on intro

Open-source \nids{}, like Snort~\cite{snort} and
Suricata~\cite{suricata}, provide rulesets for many existing
attacks. Unfortunately, attackers actively look for new mechanisms to

\begin{wrapfigure}[8]{r}{0.21\textwidth}
    \vspace{-2ex}  
    \centering
    \scalebox{0.9}{  
      \begin{tikzpicture}
        \begin{axis}[
            width = 0.3\textwidth,
            height = 0.2\textwidth,            
            bar width=3pt,
            scaled ticks=false,
            ymajorticks=false,
            ybar stacked,
            ymax=250,
            enlargelimits=0.09,
            clip=true,
            legend style={at={(0.5,-0.15)},anchor=north,legend columns=-1},          
            xtick=data,
            xticklabel style={rotate=45, font=\footnotesize},
            tick pos=left,
            nodes near coords,
            every node near coord/.append style={font=\footnotesize},
            nodes near coords align={vertical},
            symbolic x coords={Jan, Feb, Mar, Apr, May, Jun, Jul, Aug, Sep, Oct, Nov, Dec},]
            \addplot coordinates {(Jan, 114) (Feb, 155) (Mar, 117) (Apr, 159) (May, 110) (Jun, 231) (Jul, 122) (Aug, 167) (Sep, 107) (Oct, 225) (Nov, 148) (Dec, 127)};
        \end{axis}
      \end{tikzpicture}
    }
    \vspace{-5ex}
    \caption{\label{fig:distribution-rules-per-month}2019 new rules.}
\end{wrapfigure}
exploit systems through the
network. Figure~\ref{fig:distribution-rules-per-month} shows the
number of rules added per month in the year of 2019 to the ``Emerging
Threats Open Ruleset''~\cite{emerging-threats-open}, a popular public
data set of \suri\ rules. A closer observation on the changelogs of
this ruleset~\cite{emerging-threats-changelogs} shows that new rules
are often created.  Rulesets need to be continuously revised as the
community finds and advertises vulnerabilities, which eventually can
be exploited by hackers.  Unfortunately, creating these rules is
time-consuming and error-prone~\cite{vollmer-etal-cics2011,alparslan-blog-suri}. The language for
describing rules is non-trivial (see
Section~\ref{sec:example-suricata-rules}) and testing correctness of a
rule requires data and careful inspection.

@@

\tname{} takes on input one or more examples of an attack, identified by a domain expert, 
and produces on output a ranked list of rules for that attack. It uses public data
sets of benign messages~\cite{tcpreplay,stratosphere-normal} and
rulesets~\cite{emerging-threats-open} to create those rules. The
technique is organized as a pipeline of three components. First,
\tname\ uses the messages that manifest an attack to create a
potentially over-specified seed rule. Then, it uses that seed rule and
benign traffic to derive a set of plausible rules that isolate the
benign traffic from the malicious traffic. Finally, it uses existing
rules to learn heuristics to rank the plausible rules.

%\Mar{@Lucas, please look for constant
  %numbers as this *in text* and use macros instead. We do not want to
  %leave inconsistent data in the paper!}
We evaluated \tname\ against \totalAttacks attacks of various kinds, such as 
Denial-of-Service (DoS)~\cite{denial-of-service},
Active Reconnaissance (AREC)~\cite{active-reconnaissance} and Privilege
escalation (PESC)~\cite{privilege-escalation}. %\Fix{...}\rui{add few examples}. 
The observed results are encouraging. For 
example, \tname\ ranked the correct rule among the top ten rules in 
\percTopFiveRanking\ of the cases. \rui{should we quote the same numbers as in the abstract?}

>>>>

\begin{wrapfigure}[7]{r}{0.21\textwidth}
    \vspace{-4ex}  
    \centering
    \scalebox{0.9}{  
      \begin{tikzpicture}
        \begin{axis}[
            width = 0.3\textwidth,
            height = 0.2\textwidth,            
            bar width=3pt,
            scaled ticks=false,
            ymajorticks=false,
            ybar stacked,
            ymax=250,
            enlargelimits=0.09,
            clip=true,
            legend style={at={(0.5,-0.15)},anchor=north,legend columns=-1},          
            xtick=data,
            xticklabel style={rotate=45, font=\footnotesize},
            tick pos=left,
            nodes near coords,
            every node near coord/.append style={font=\footnotesize},
            nodes near coords align={vertical},
            symbolic x coords={Jan, Feb, Mar, Apr, May, Jun, Jul, Aug, Sep, Oct, Nov, Dec},]
            \addplot coordinates {(Jan, 114) (Feb, 155) (Mar, 117) (Apr, 159) (May, 110) (Jun, 231) (Jul, 122) (Aug, 167) (Sep, 107) (Oct, 225) (Nov, 148) (Dec, 127)};
        \end{axis}
      \end{tikzpicture}
    }
    \vspace{-5ex}
    \caption{\label{fig:distribution-rules-per-month}2019 new rules.}
\end{wrapfigure}
Open-source \nids{}, like Snort~\cite{snort} and
Suricata~\cite{suricata}, provide rulesets for many existing
attacks. Unfortunately, attackers actively look for new mechanisms to
exploit systems through the
network. Figure~\ref{fig:distribution-rules-per-month} shows the
number of rules added per month in the year of 2019 to the ``Emerging
Threats Open Ruleset''~\cite{emerging-threats-open}, a popular public
data set of \suri\ rules. A closer observation on the changelogs of
this ruleset~\cite{emerging-threats-changelogs} shows that new rules
are often created.  Rulesets need to be continuously revised as the
community finds and advertises vulnerabilities, which eventually can
be exploited by hackers.  Unfortunately, creating these rules is
time-consuming and error-prone~\cite{vollmer-etal-cics2011,alparslan-blog-suri}. The language for
describing rules is non-trivial (see
Section~\ref{sec:example-suricata-rules}) and testing correctness of a
rule requires data and careful inspection.

@@

%\rui{We ought to explain how does one get hold on examples. How diffcult are they to obtain?} \Gui{Given a captured network traffic, an expert must identify which packets can be used to generate the rule. These packets are separated and given as input to Syrius.}

\tname{} takes on input one or more examples of an attack, identified by a domain expert, and produces
on output a ranked list of rules for that attack. One assumption that
we make is that system administrators are able to capture the traffic
that perpetrated the attack. The capture can be obtained by other
institutions, who shared that information, or the organization itself
that suffered the attack and wants to avoid that
repeated. \tname\ uses public data sets of benign
messages~\cite{tcpreplay,stratosphere-normal} and
rulesets~\cite{emerging-threats-open} to create those rules. The
technique is organized as a pipeline of three components. First,
\tname\ uses the messages that manifest an attack to create a
potentially over-specified seed rule. Then, it uses that seed rule and
benign traffic to derive a set of plausible rules that isolate the
benign traffic from the malicious traffic. Finally, it uses existing
rules to learn heuristics to rank the plausible rules.

%\Mar{@Lucas, please look for constant
  %numbers as this *in text* and use macros instead. We do not want to
  %leave inconsistent data in the paper!}

We evaluated \tname\ against \totalAttacks attacks of various kinds
(\eg{} Denial-of-Service (DoS)~\cite{denial-of-service}, Active
Reconnaissance (AREC)~\cite{active-reconnaissance} and Privilege
escalation (PESC)~\cite{privilege-escalation}) and from various
sources. Preliminary results provide early evidence that \tname\ is
promising. It reported the ground truth rule in \emph{all} cases and
reported that rule within the top-1, top-3, top-5, and top-10 rules of
the ranking, respectively, in \percTopOneRanking\,
\percTopThreeRanking, \percTopFiveRanking, and \percTopTenRanking\ of
the cases.
